Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-09-15 paper/verification-conditions.tex
Revision to make paper flow with overview

The AGREE specification language, and semantics, are inspired by the
Lustre language \cite{10.1145/41625.41641}. Lustre semantics are
synchronous data-flow where the inputs and outputs of components are
streams.  Informally, streams assign values to expressions at each
point in time.  AGREE contracts express relationships between input
and output streams by assuming a set of allowed input streams and then
guaranteeing a corresponding set of allowed output streams.

\subsection{Components}
A system implementation is described by its ports, its subcomponents,
and the connections between ports on different subcomponents.  Ports
in AGREE are inherited from AADL.  There are three kinds of port:
\emph{event} ports, \emph{data} ports, and \emph{event data} ports.
An event port carries no data and only signals occurrence (or not) of the
event. A data port is a sampled port: the data can be read at anytime.
An event data port indicates when data is present with an event.

@@

From the system and component contracts, AGREE generates a set of
verification conditions to show that a system's component
implementation is correct~\cite{agree2013}.  The AGREE model checker
is then invoked to prove or disprove the verification
conditions. The contracts and verification conditions are expressed in
\emph{past-time linear temporal logic} (PLTL) \cite{10.1093/jigpal/8.1.55}.

@@

The two PLTL operators necessary for the AGREE generated verification
conditions are $\globally$ that looks forward in time along
the trace and $\historically$ (historically) that looks backward in
time along the trace.  These are defined as
\begin{eqnarray*}
 (\pi, i) \models \globally(f) & \iff & \forall j \ge i, (\pi, j) \models f \\
(\pi, i) \models \historically(f) & \iff & \forall 0 \le j \le i, (\pi, j) \models f
\end{eqnarray*}
The $\models$-operator is read as \emph{satisfies}.  A trace at a
moment in time satisfies $\globally(f)$ if and only if it satisfies
$f$ in the current and all future states of $\pi$.  $\globally(f)$ is
invariant from the current moment into the future and $\historically$ is
invariant from the beginning of the trace to the current moment.

@@

If all the verification conditions hold (AGREE uses $k$-inductive
model checking to automatically prove or disprove each generated
verification condition), then the system is said to be \emph{correct},
meaning that the system composition meets input assumptions at each
input as well as the guarantees on the system output. A consequence of
this result is that $\globally(\historically(\assumes) \implies
\guarantees)$ holds for the system contract.  Therefore the Liskov
substitution principle applies and the implementation is a safe
substitution for the system contract \cite{10.1145/62139.62141}.

The expanded property lists in \figref{fig:example-certificate} and
\figref{fig:hardened-certificate} are the results from verifying or
disproving the above verification conditions.  The additional
unexpanded results at the bottom of the figures prove
\emph{self-consistency} in the contracts.
A contract is self-consistent if it does not guarantee two different values in the same moment of time, \ie, the contract is not self-contradicting.
AGREE generates additional verification
conditions that prove each component contract, and the composition of
contracts, self-consistent.

>>>>

The \agr\ specification language, and semantics, are inspired by the
Lustre language \cite{10.1145/41625.41641}. Lustre semantics are
synchronous data-flow where the inputs and outputs of components are
streams.  Informally, streams assign values to expressions at each
point in time.  \agr\ contracts express relationships between input
and output streams by assuming a set of allowed input streams and then
guaranteeing a corresponding set of allowed output streams.

\subsection{Components}
A system implementation is described by its ports, its subcomponents,
and the connections between ports on different subcomponents.  Ports
in \agr\ are inherited from AADL.  There are three kinds of port:
\emph{event} ports, \emph{data} ports, and \emph{event data} ports.
An event port carries no data and only signals occurrence (or not) of the
event. A data port is a sampled port: the data can be read at anytime.
An event data port indicates when data is present with an event.

@@

From the system and component contracts, \agr\ generates a set of
verification conditions to show that a system's component
implementation is correct~\cite{agree2013}.  The \agr\ model checker
is then invoked to prove or disprove the verification
conditions. The contracts and verification conditions are expressed in
\emph{past-time linear temporal logic} (PLTL) \cite{10.1093/jigpal/8.1.55}.

@@

The two PLTL operators necessary for the \agr\ generated verification
conditions are $\globally$ that looks forward in time along
the trace and $\historically$ (historically) that looks backward in
time along the trace.  These are defined as
\begin{eqnarray*}
 (\pi, i) \models \globally(f) & \iff & \forall j \ge i, (\pi, j) \models f \\
(\pi, i) \models \historically(f) & \iff & \forall 0 \le j \le i, (\pi, j) \models f
\end{eqnarray*}
The $\models$-operator is read as \emph{satisfies}.  A trace at a
moment in time satisfies $\globally(f)$ if and only if it satisfies
$f$ in the current and all future states of $\pi$.  $\globally(f)$ is
invariant from the current moment into the future and $\historically$ is
invariant from the beginning of the trace to the current moment.

@@

If all the verification conditions hold (\agr\ uses $k$-inductive
model checking to automatically prove or disprove each generated
verification condition), then the system is said to be \emph{correct},
meaning that the system composition meets input assumptions at each
input as well as the guarantees on the system output. A consequence of
this result is that $\globally(\historically(\assumes) \implies
\guarantees)$ holds for the system contract.  Therefore the Liskov
substitution principle applies and the implementation is a safe
substitution for the system contract \cite{10.1145/62139.62141}.

The expanded property lists in \figref{fig:example-certificate} and
\figref{fig:hardened-certificate} are the results from verifying or
disproving the above verification conditions.  The additional
unexpanded results at the bottom of the figures prove
\emph{self-consistency} in the contracts.
A contract is self-consistent if it does not guarantee two different values in the same moment of time, \ie, the contract is not self-contradicting.
\agr\ generates additional verification
conditions that prove each component contract, and the composition of
contracts, self-consistent.
